\chapter{Introduction}
\label{ch:into} % This how you label a chapter and the key (e.g., ch:into) will be used to refer this chapter ``Introduction'' later in the report. 
% the key ``ch:into'' can be used with command \ref{ch:intor} to refere this Chapter.

% \textbf{Guidance on introduction chapter writing:} Introductions are written in the following parts:
% \begin{itemize}
%     \item A brief  description of the investigated problem.
%     \item A summary of the scope and context of the project, i.e., what is the background of the topic/problem/application/system/algorithm/experiment/research question/hypothesis/etc. under investigation/implementation/development [whichever is applicable to your project].
%     \item The aims and objectives of the project.
%     \item A description of the problem and the methodological approach adopted to solve the problem.
%     \item A summary of the most significant outcomes and their interpretations.
%     \item Organization of the report. 
% \end{itemize}


% Consult \textbf{your supervisor} to check the content of the introduction chapter. In this template, we only offer basic sections of an introduction chapter. It may  not be complete and comprehensive. Writing a report is a subjective matter, and a report's style and structure depend on the ``type of project'' as well as an individual's preference. This template suits the following project paradigms:
% \begin{enumerate}
%     \item software engineering and software/web application development;
%     \item algorithm implementation, analysis and/or application;  
%     \item science lab (experiment); and
%     \item pure theoretical development (not mention extensively).
% \end{enumerate}

% Use only a single \textbf{font} for the body text. We recommend using a clean and electronic document friendly font like \textbf{Arial} or \textbf{Calibri} for MS-word (If you create a report in MS word). If you use this template, DO NOT ALTER the template's default font ``amsfont default computer modern''. The default \LaTeX~font ``computer modern'' is also acceptable. 

% The recommended body text \textbf{font size} is minimum \textbf{11pt} and minimum one-half line spacing. The recommended figure/table caption font size is minimum 10pt. The footnote\footnote{Example footnote: footnotes are useful for adding external sources such as links as well as extra information on a topic or word or sentence. Use command \textbackslash footnote\{...\} next to a word to generate a footnote in \LaTeX.} font size is minimum 8pt. DO NOT ALTER the font setting of this template.   

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
\label{sec:into_back}
% Describe to a reader the context of your project. That is, what is your project and what its motivation. Briefly explain the major theories, applications, and/or products/systems/algorithms whichever is relevant to your project.



The proficiency of Large Language Models (LLMs) like ChatGPT in generating working code is 
intrinsically linked to their understanding and application of data structures and algorithms 
(DSA). These fundamental components form the backbone of efficient and effective software 
development. DSA are critical in determining how data is organized, stored, and manipulated 
within a program, impacting everything from the execution speed to resource utilization. The 
ability of an AI model to adeptly handle these aspects is indicative of its depth in coding 
knowledge and its applicability in real-world software development scenarios.

In this context, the capabilities of ChatGPT in DSA are particularly noteworthy. This model has 
demonstrated a remarkable ability to not only understand and implement standard data structures 
and algorithms but also to apply them creatively to solve complex problems. The implication of 
this proficiency is significant; it suggests that LLMs like ChatGPT can be invaluable tools in 
the software development process, assisting in everything from initial problem analysis to the 
formulation of efficient algorithmic solutions.

Furthermore, the application of DSA in code generation by LLMs also opens up possibilities for 
more advanced software development applications. For instance, an AI model proficient in DSA 
can potentially assist in optimizing existing codebases, refactoring inefficient structures, or 
even suggesting algorithmic improvements. 

In the current landscape of AI-driven code generation, particularly with models like ChatGPT, 
there is a noticeable challenge in consistently generating code that is both efficient and 
correct. The generation of correct code is fundamental, as errors in code can lead to 
significant issues in software development, ranging from minor bugs to critical system 
failures. While ChatGPT exhibits a strong capacity for code generation especially in DSA, as 
evidenced in the work by \cite{arefin2023unmasking}, its performance can vary, especially in 
terms of efficiency and adherence to best practices in DSA. This variability underscores the 
necessity for more refined techniques in interacting with these models to elicit the highest 
quality of code output.

Prompt engineering emerges as a crucial element in this context. It plays a critical role in 
harnessing their full potential, especially in code generation. It involves meticulously 
crafting input prompts to effectively communicate the requirements and constraints of a given 
problem to the AI model. This practice is particularly essential in programming scenarios where 
the quality of output is directly influenced by how the problem is presented to the model, 
i.e., the clarity and specificity of instructions can significantly influence the accuracy and 
utility of the generated code. A well-engineered prompt can lead to solutions that are not only 
syntactically and logically correct but also optimized in terms of performance and resource 
management. The significance of prompt design in obtaining optimal results from generative 
tasks has been highlighted in the research by \cite{chen2021evaluating} on GPT models.

The implications of a system proficient in accurately generating working, logical code are far-
reaching. In the future, such capabilities could revolutionize software development, enabling 
faster deployment of robust and sophisticated applications. Furthermore, as noted by \cite{chen2021evaluating} in their analysis of code generation models, these advancements could democratize 
programming, allowing individuals with limited coding expertise to develop software solutions 
through intuitive, natural language instructions. This could potentially lead to a surge in 
innovation, as barriers to software creation are lowered, and a broader range of perspectives 
are brought into the technology development process.

The current research aims to delve deeper into ChatGPT’s capabilities in DSA. It aims to assess 
and quantify the impact of prompt engineering on ChatGPT’s performance in generating 
algorithmic solutions. By evaluating the model’s performance in this domain, the study seeks to 
shed light on the extent to which ChatGPT can accurately and effectively generate code 
solutions that are not just correct in their logic, but also optimal in their use of data 
structures and algorithms. This exploration is pivotal, as it directly relates to the practical 
utility of LLMs in software engineering, a field where efficiency and optimization are 
paramount.




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problem statement}
\label{sec:intro_prob_art}
% This section describes the investigated problem in detail. You can also have a separate chapter on ``Problem articulation.''  For some projects, you may have a section like ``Research question(s)'' or ``Research Hypothesis'' instead of a section on ``Problem statement.'
The central problem this study addresses is the assessment of the impact of prompt engineering 
on the performance of ChatGPT, specifically in solving undergraduate-level data structures and 
algorithms (DSA) questions. The effectiveness of Large Language Models (LLMs) like ChatGPT in 
code generation has been increasingly recognized. However, their ability to consistently 
produce efficient and correct solutions in the context of complex DSA problems remains an area 
requiring deeper exploration. This research hypothesizes that through strategic prompt 
engineering, the proficiency of ChatGPT in generating accurate and optimized solutions to DSA 
challenges can be significantly enhanced.

Prompt engineering, in this context, refers to the deliberate design and structuring of input 
prompts to guide ChatGPT in understanding and effectively responding to the intricacies of DSA 
problems. The hypothesis is grounded in the premise that the manner in which a problem is 
presented to an LLM can profoundly influence the quality of the generated solution. This 
hypothesis aligns with the findings of \cite{radford2019language}, who highlighted the 
importance of prompt design in achieving desired outcomes from generative AI models. The study 
aims to empirically test this hypothesis by systematically varying the prompts used to interact 
with ChatGPT and evaluating the resulting code's correctness, efficiency, and adherence to DSA 
best practices.

The significance of this investigation lies not only in its potential to enhance the 
understanding of prompt engineering as a tool for optimizing LLM performance but also in its 
broader implications for the field of AI-driven software development. By establishing a clear 
correlation between prompt engineering and the quality of ChatGPT-generated code, the study 
seeks to contribute to the development of more effective methodologies for leveraging LLMs in 
technical and educational settings.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Aims and objectives}
\label{sec:intro_aims_obj}
% Describe the ``aims and objectives'' of your project. 

% \textbf{Aims:} The aims tell a read what you want/hope to achieve at the end of the project. The  aims define your intent/purpose in general terms.  

% \textbf{Objectives:} The objectives are a set of tasks you would perform in order to achieve the defined aims. The objective statements have to be specific and measurable through the results and outcome of the project.
The overarching aim of this project is to conduct a comprehensive evaluation of ChatGPT's 
capabilities in solving data structures and algorithms problems typically encountered at the 
undergraduate level. This aim encompasses several specific objectives:

\textbf{Development of Tailored Prompts}: One of the primary objectives is to develop tailored 
prompts that enhance ChatGPT’s problem-solving skills in data structures and algorithms. This 
involves designing prompts that not only accurately describe the problem but also guide the 
model towards optimal problem-solving approaches. The effectiveness of these prompts will be 
gauged by their ability to elicit accurate, efficient, and sophisticated solutions from ChatGPT.

\textbf{Curating a Comprehensive Standard Test Set}: The project will involve curating a 
comprehensive set of standard undergraduate-level questions in data structures and algorithms. 
The selected test set for this research includes a combination of Easy, Medium and Hard questions which reflect a standard DSA question set an undergraduate student might encounter in a real world scenario.

\textbf{Automation of the Evaluation Process}: Automating the evaluation of ChatGPT’s responses 
is another critical objective. This will be achieved by integrating ChatGPT with the LeetCode 
API (graphql), allowing for an efficient and systematic assessment of the solutions' correctness and 
practical applicability. Automation will ensure consistency in evaluation and enable the 
handling of a large volume of test cases.

\textbf{Comparison of Effectiveness}: The project aims to compare the effectiveness of 
ChatGPT's responses with and without prompt engineering. This comparative analysis will help 
ascertain the impact of prompt engineering on the model's performance. It will involve 
analyzing the solutions generated by ChatGPT under different prompting conditions to evaluate 
how variations in prompt design influence the accuracy and quality of the generated code.

Through these objectives, the project seeks to provide a detailed analysis of how effectively 
ChatGPT can be utilized in the domain of data structures and algorithms, and the extent to 
which prompt engineering can enhance its performance. The results are expected to offer 
valuable insights into the practical applications of LLMs in technical education and software 
development.









%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Solution approach}
\label{sec:intro_sol} % label of Org section
% Briefly describe the solution approach and the methodology applied in solving the set aims and objectives.

% Depending on the project, you may like to alter the ``heading'' of this section. Check with you supervisor. Also, check what subsection or any other section that can be added in or removed from this template.
The methodology of this project is structured around a blend of prompt engineering and 
automated evaluation techniques, designed to rigorously test ChatGPT's proficiency in solving 
data structures and algorithms problems.

By employing these methodical and automated approaches outlined below, the study aims to 
provide a detailed analysis of ChatGPT's problem-solving abilities in the context of data 
structures and algorithms. The findings are expected to offer significant insights into the 
potential of LLMs in technical education and software development, particularly in the realm of 
algorithmic thinking and coding proficiency.

\subsection{Automated System Development and Integration}
\label{sec:Leetcode}
% You may or may not need subsections here. Depending on your project's needs, add two or more subsection(s). A section takes at least two subsections. 
A key component of our methodology involves developing an automated system that interfaces with 
both the ChatGPT and LeetCode APIs. This system will automatically fetch questions from 
LeetCode, using graphql API behind leetcode.com. These questions, 
representative of common undergraduate-level data structures and algorithms challenges, form 
the basis of our evaluation set.

\subsection{Solution Generation and Submission}
\label{sec:intro_some_sub2}
% Depending on your project's needs, add more section(s) and subsection(s).
Once a question is fetched, it is passed to the ChatGPT API to generate a solution. This 
process is repeated multiple times for each question, with each iteration using a different set 
of prompts. These prompts are strategically designed to guide ChatGPT towards generating more 
effective solutions. Each generated solution is then submitted back to LeetCode through its 
API, enabling us to assess its correctness and efficiency.

\subsection{Evaluation Metrics}
\label{sec:intro_some_subsub1}
% The command \textbackslash subsubsection\{\} creates a paragraph heading in \LaTeX.
The primary metrics for evaluating each solution are the number of test cases passed and the 
run time. The number of test cases passed is a direct indicator of the solution's correctness, 
reflecting how well the code meets the problem's requirements. The run time, on the other hand, 
provides insight into the efficiency and optimization of the solution. Together, these metrics 
are critical for evaluating the effectiveness of different prompts and the overall capability 
of ChatGPT in solving complex DSA questions.

\subsection{Comprehensive Testing Across Standard Test Questions}
\label{sec:intro_some_subsub2}
% Write your text here...
This entire process, from question retrieval to solution evaluation, is systematically repeated 
for all the standard test questions outlined in the Aims \& Objectives section. This 
comprehensive approach ensures a thorough assessment of ChatGPT’s capabilities and the impact 
of various prompt engineering strategies on its performance.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Summary of contributions and achievements} %  use this section 
% \label{sec:intro_sum_results} % label of summary of results
% Describe clearly what you have done/created/achieved and what the major results and their implications are. 


% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Organization of the report} %  use this section
% \label{sec:intro_org} % label of Org section
% Describe the outline of the rest of the report here. Let the reader know what to expect ahead 
% in the report. Describe how you have organized your report. 

% \textbf{Example: how to refer a chapter, section, subsection}. This report is organised into seven chapters. Chapter~\ref{ch:lit_rev} details the literature review of this project. In Section~\ref{ch:method}...  % and so on.

% \textbf{Note:}  Take care of the word like ``Chapter,'' ``Section,'' ``Figure'' etc. before the \LaTeX command \textbackslash ref\{\}. Otherwise, a  sentence will be confusing. For example, 
% In \ref{ch:lit_rev} literature review is described. In this sentence, the word ``Chapter'' is 
% missing. Therefore, a reader would not know whether 2 is for a Chapter or a Section or a Figure.

\section{Summary of Contributions and Achievements}
\label{sec:intro_sum_results}
This project has achieved significant milestones in evaluating the capabilities of ChatGPT in solving complex data structures and algorithms (DSA) problems through an automated testing framework. Major contributions of this work include:

\begin{itemize}
    \item \textbf{Development of an Automated System:} A fully automated pipeline was created for fetching coding problems, generating solutions using ChatGPT, and evaluating these solutions against standardized test cases. This system is scalable and can be adapted for broader AI testing applications.
    \item \textbf{Role-Based Prompt Engineering:} This work extensively explored how different prompts influence the performance of ChatGPT. Through careful design of role-based prompts, it was possible to direct the AI to approach problems with varying strategies, significantly affecting the quality of the solutions generated.
    \item \textbf{Comprehensive Testing and Analysis:} A vast array of DSA problems was used to test the model, providing a deep insight into its strengths and weaknesses across different topics. The findings reveal that while ChatGPT excels in areas requiring mathematical reasoning, it struggles with tasks that need complex data structure manipulation.
    \item \textbf{Identification of Key Patterns:} The project identified critical patterns in errors and solution quality, which can help refine the AI model's training and prompt engineering for improved performance.
\end{itemize}

The implications of these results are vast, providing a foundation for future research in AI-driven coding solutions and enhancing the understanding of how language models can be effectively utilized in programming education and professional software development.

\section{Organization of the Report}
\label{sec:intro_org}
This report is organized into seven main chapters, each dedicated to a different aspect of the research, ensuring a logical flow and comprehensive understanding of the project.

\begin{itemize}
    \item \textbf{Chapter~\ref{ch:into}: Introduction} - This chapter sets the stage for the report, outlining the research problem, objectives, and the scope of the study.
    \item \textbf{Chapter~\ref{ch:lit_rev}: Literature Review} - Details the existing research on AI capabilities in coding, specifically focusing on the performance of language models like ChatGPT in technical domains.
    \item \textbf{Chapter~\ref{ch:method}: Methodology} - Describes the methods used for data collection, prompt engineering, solution generation and submission, and the automated system development and integration.
    \item \textbf{Chapter~\ref{ch:results}: Results} - Presents the findings from the extensive testing of ChatGPT, analyzing its performance across various prompts and topics.
    \item \textbf{Chapter~\ref{ch:discussion}: Discussion and Analysis} - Discusses the implications of the results, the effectiveness of different prompts, and the model's performance on various DSA topics.
    \item \textbf{Chapter~\ref{ch:conclusions}: Conclusions and Future Work} - Summarizes the key findings and discusses potential areas for future research.
    \item \textbf{Chapter~\ref{ch:reflection}: Reflection} - Provides a personal account of the research experience, reflecting on the learning and challenges encountered during the project.
\end{itemize}

Appendices and additional supporting material can be found at the end of the report, providing further details on data and methodologies used throughout the research.

